Quantitative processing of broadband data as implemented in a scientific splitbeam echosounder

### 2022-04-26 Paper/Manuscript.tex
Edits to Discussion & tweaks to rest of doc.

\subsection{Accompanying code}
The code accompanying this paper is written in the Python programming language and is available in the supplementary materials. All processing steps with respective figures in the paper can be reproduced by running the main scripts (\code{TSf.py} and \code{Svf.py}). Without loss of generality, we use the Simrad EK80 echosounder as an example since it is currently the most commonly used broadband system in the marine ecosystem acoustics field.

@@

As a metallic sphere is a rather simple and near-ideal scatterer, we use $\sv$ from a school of non-swimbladdered fish (fig.~\ref{Fig_Sv_echogram}) collected with a 120 kHz centre frequency transducer to illustrate the processes to obtain $\sv(m)$.

@@

The normalized discrete Fourier transform of the windowed data, $\ypcvolumenormf(\samplesymf)$, is then given by:
%
\begin{equation}
\label{eq:FFT_volume_norm}
\ypcvolumenormf(\samplesymf) = \frac{\ypcvolumef(\samplesymf)}{\ymfautof(\samplesymf)},
\end{equation}
%
and received power into a matched load, $\prxevf(\samplesymf)$, is estimated from:
%
\begin{equation}
\label{eq:prx_FFT_volume}
\prxevf(\samplesymf) = \nchannels \left( \frac{|\ypcvolumenormf(\samplesymf)|}{2\sqrt{2}} \right)^2 \left( \frac{|\zrxe+\ztde|}{|\zrxe|}\right)^2 \\
\frac{1}{|\ztde|}. % All channels (*4), matched load (/2), and effective values (/sqrt(2))
\end{equation}
%
Finally, the discretized estimate of $\sv(\freqsym)$, $\sv(\samplesymf)$, is given by:

%
\begin{equation}
\label{eq:Sv_FFT}
\sv(\freqsym) = 10\log_{10}(\prxevf(\samplesymf)) + 2\absorp(\freqsym) \range_c - 10\log_{10}\left( \frac{\ptxe \wlen^2 \cw \tslide \eqang(\freqsym) \gainzero^2(\freqsym) }{32\pi^2} \right).
\end{equation}
where the sample index $\samplesymf$ corresponding to frequency $\freqsym$ can be estimated using
\begin{equation}
\label{eq:m(f)}
\samplesymf = \lfloor \frac{\freqsym}{\fsdec} \ndft \rfloor \mod \ndft
\end{equation}

By selecting a set of centre samples $\timesym$, $\sv$ values can be presented as a function of range ($\samplesymt$) and frequency ($freqsym$) for each ping. The range for centre samples $\samplesymt$ could be chosen to half the window length or any other grid that the user prefer the data presented to be in. This can be useful when combining the $\sv(\freqsym)$ across a range of transducers. In our example we have simply chosen n to be on the same order of the range resolution (fig.~\ref{Fig_Sv_m_n}).

@@

For acoustic abundance estimation and classification purposes it is common to integrate the $\sv$ over a range (15 to 34 m in the example, covering a school of non-swimbladdered fish, Fig.~\ref{Fig_Sv_echogram}). It is normal to average $\sv$ over several pings to obtain an unbiased estimate, but here just one ping is used for illustrative purposes (fig.~\ref{Fig_Sv_avg}). Even though this is for a single ping it is still possible to observe a positive slope of the frequency response that is indicative of non-swimbladdered fish.

@@

The trend for increasing $\sv$ with frequency is well-known \citep{korneliussen2010} and is consistent with the trend observed in this example. In contrast to data from isolated scatterers, such as metallic spheres, the benefit of pulse compression on the backscatter from an object that generates many overlapping echoes is not immediately obvious (fig.~\ref{Fig_Sv_echogram}).

@@

The use of broadband signals in fisheries acoustics is a developing field, and our contribution represents a comprehensive description of data processing steps. The inclusion of a corresponding code ensures a complete description. One could also envision that signal processing steps designed for specific purposes beyond these standard parameters can be developed, and our code may then serve as a starting point for further development. We also envision that the paper and code will be used both for educational purposes and for understanding the standard signal processing.

The contribution includes all the minor details required for the processing steps to work in practice. This includes steps well founded in the literature as well as practical and more ad-hoc choices, and a few examples of such choices follows. How to handle gaps in the calibration data, the choice of transmit pulse including tapering, calculation of efficient pulse duration, and decimation factors and filtering are examples of such choices. When convolving the received signal with the transmit pulse, there are various approaches to handle edge cases, and, in our implementation, we simply chose to remove the edge cases. We have also assumed a four-sector transducer, and the code must be adapted to other beam configurations if other configurations are needed. When estimating $\ts(\freqsym)$ and $\sv(\freqsym)$ the resolution and accuracy will depend on the length $\ndft$ of the Fourier transform and the actual choice will be a compromise between accuracy and computational speed. Our objective is not to provide an evaluation of all these choices, but to serve as a reference implementation when benchmarking the choices.

$\ts(\freqsym)$ is a common metric for studying single targets, used to extract features for single individuals. These features includes size, target classification and is used to extract behaviour through tracking. In our implementation, the target strength by frequency calculation assumes that a single target has successfully been identified. For this to work in proactive a robust single target detector is required. Several single echo detection algorithms exist, and various single echo detection algorithms may be required depending on situation. Typical SED algorithms are based on traditional single frequency pulses, and utilizing the information in the broadband data improved SEDs may be envisioned, but this is outside the scope of this paper.

$\sv(\freqsym)$ is a key parameter for echo integration. To estimate $\sv(\freqsym)$ a Fourier transform is used, repeatedly applied via a sliding window in range. The recommended size of the window is two times the pulse length, and is chosen as a compromise between spatial and frequency resolution. Since the duration of the sliding window can cause the spreading loss compensation to differ between the beginning and end of the window, the compensation for spreading loss is performed on the pulse compressed time domain data before the transform. Absorption loss compensation is also range dependent (and frequency dependent), but is insignificant for typical operating frequencies at short ranges and between the beginning and the end of the window at longer ranges. The compensation for absorption loss is therefore performed after applying the discrete Fourier transform. The choice of window also allow the data to be cast on a predefined range-frequency grid, which can be used to fit data across transducers to an n-dimensional tensor typically employed by machine learning frameworks like Keras, Pytorch, and Tensor flow.

The formulation presented in this paper results in several frequency dependent parameters, such as transducer gain, two-way equivalent beam angle, and the water absorption coefficient, that are required to quantitatively estimate $\ts(\freqsym)$ and $\sv(\freqsym)$ from received broadband signals. Methods to estimate these are not within the scope of this paper, but common practise is to use the conventional sphere backscatter calibration methodology \citep{demerCalibrationAcousticInstruments2015} slightly enhanced for broadband \citep{hobaekCharacterizationTargetSpheres2013,lavery2017}. We note that these methods do not provide an operational method to estimate $\teff$ or $\Phi(f)$, especially for ship-mounted transducers, and that empirical measurements of these parameters are necessary to fully calibrate both narrowband and broadband echosounders.

A set of equations for calculating calibrated, frequency-dependent, target strength and volume backscatter from broadband echosounder signals have been presented. Providing an open-source resource for anyone interested in learning and developing the techniques further. We note that the processing equations and methodology presented in this paper are similar to version 1.12.4 and earlier of the Simrad EK80 software.

>>>>

\subsection{Accompanying code}
The code accompanying this paper is written in the Python programming language and is available in the supplementary materials. All processing steps with respective figures in the paper can be reproduced by running the main scripts (\code{TSf.py} and \code{Svf.py}). Without loss of generality, we use the \ek{} echosounder as an example since it is currently the most commonly used broadband system in the marine ecosystem acoustics field.

@@

As a metallic sphere is a rather simple and near-ideal scatterer, we use $\sv$ from a school of non-swimbladdered fish (Fig.~\ref{Fig_Sv_echogram}) collected with a 120 kHz centre frequency transducer to illustrate the processes to obtain $\sv(m)$.

@@

The normalized discrete Fourier transform of the windowed data, $\ypcvolumenormf(\samplesymf)$, is then given by:
%
\begin{equation}
\label{eq:FFT_volume_norm}
\ypcvolumenormf(\samplesymf) = \frac{\ypcvolumef(\samplesymf)}{\ymfautof(\samplesymf)},
\end{equation}
%
and received power into a matched load, $\prxevf(\samplesymf)$, is estimated from:
%
\begin{equation}
\label{eq:prx_FFT_volume}
\prxevf(\samplesymf) = \nchannels \left( \frac{|\ypcvolumenormf(\samplesymf)|}{2\sqrt{2}} \right)^2 \left( \frac{|\zrxe+\ztde|}{|\zrxe|}\right)^2 \\
\frac{1}{|\ztde|}. % All channels (*4), matched load (/2), and effective values (/sqrt(2))
\end{equation}
%
Finally, the discretized estimate of $\sv(\freqsym)$, $\sv(\samplesymf)$, is given by:
%
\begin{equation}
\label{eq:Sv_FFT}
\sv(\freqsym) = 10\log_{10}(\prxevf(\samplesymf)) + 2\absorp(\freqsym) \range_c - 10\log_{10}\left( \frac{\ptxe \wlen^2 \cw \tslide \eqang(\freqsym) \gainzero^2(\freqsym) }{32\pi^2} \right).
\end{equation}
where the sample index $\samplesymf$ corresponding to frequency $\freqsym$ can be estimated using Eq. {\ref{eq:m(f)}}.

By selecting a set of centre samples $\timesym$, $\sv$ values can be presented as a function of range ($\samplesymt$) and frequency ($freqsym$) for each ping. The range for centre samples $\samplesymt$ could be chosen to half the window length or any other grid that the user prefer the data presented to be in. This can be useful when combining the $\sv(\freqsym)$ across a range of transducers. In our example we have simply chosen n to be on the same order of the range resolution (Fig.~\ref{Fig_Sv_m_n}).

@@

For acoustic abundance estimation and classification purposes it is common to integrate the $\sv$ over a range (15 to 34 m in the example, covering a school of non-swimbladdered fish, Fig.~\ref{Fig_Sv_echogram}). It is normal to average $\sv$ over several pings to obtain an unbiased estimate, but here just one ping is used for illustrative purposes (Fig.~\ref{Fig_Sv_avg}). Even though this is for a single ping it is still possible to observe a positive slope of the frequency response that is indicative of non-swimbladdered fish.

@@

The trend for increasing $\sv$ with frequency is well-known \citep{korneliussen2010} and is consistent with the trend observed in this example. In contrast to data from isolated scatterers, such as metallic spheres, the benefit of pulse compression on the backscatter from an object that generates many overlapping echoes is not immediately obvious (Fig.~\ref{Fig_Sv_echogram}).

@@

The use of broadband signals in fisheries acoustics is a developing field, and our contribution represents a comprehensive description of data processing steps. The inclusion of corresponding computer code ensures a more complete description. One could also envision that signal processing steps designed for specific purposes beyond these standard parameters can be developed, and our code can serve as a starting point for this. We also envision that the paper and code be used both for educational purposes and for understanding the standard broadband signal processing.

The contribution includes all the minor details required for the processing steps to work in practice. This includes steps well founded in the literature as well as practical and more ad-hoc choices, and a few examples of such choices follows. Handling of gaps in the calibration data, the choice of transmit pulse including tapering, calculation of efficient pulse duration, and decimation factors and filtering are examples of such choices. When convolving the received signal with the transmit pulse, there are various approaches to handle edge cases and in our implementation we chose to remove the edge cases. We have also assumed a four-sector transducer, and the code must be adapted to other beam configurations if other configurations are needed. When estimating $\ts(\freqsym)$ and $\sv(\freqsym)$ the resolution and accuracy will depend on the length, $\ndft$, of the Fourier transform and the actual choice will be a compromise between accuracy and computational speed. Our objective is not to provide an evaluation of all these choices, but to serve as a reference implementation when benchmarking the choices.

$\ts(\freqsym)$ is a common metric for studying single targets, used to extract features for single individuals. These features includes size, target classification and can be used to extract behaviour through tracking. In our implementation, the target strength by frequency calculation assumes that a single target has successfully been identified. For this to work well in practise, a robust single target detector is required. Several SED algorithms exist and differing such algorithms may be required depending on the situation. Typical SED algorithms are based on traditional single frequency pulses and by utilizing the additional information in broadband echoes improved SEDs may be envisioned, but this is outside the scope of this paper. One existing improvement is that broadband pulse compression is itself a type of single target detector.

$\sv(\freqsym)$ is a key parameter for echo integration. To estimate $\sv(\freqsym)$ a Fourier transform is used, repeatedly applied via a sliding window in range. The recommended size of the window is two times the pulse length, and is chosen as a compromise between spatial and frequency resolution. Since the duration of the sliding window can cause the spreading loss compensation to differ between the beginning and end of the window, the compensation for spreading loss is performed on the pulse compressed time domain data before the transform. Absorption loss compensation is also range dependent (and frequency dependent), but is insignificant for typical marine ecosystem echosounder operating frequencies at short ranges and between the beginning and the end of the window at longer ranges. The compensation for absorption loss is therefore performed after applying the discrete Fourier transform. The choice of window also allows the data to be split onto a predefined range-frequency grid, which can then be used to fit data across transducers to an n-dimensional tensor typically employed by machine learning frameworks like Keras, Pytorch, and Tensor flow.

The formulation presented in this paper requires several frequency dependent parameters, such as transducer gain, two-way equivalent beam angle, and the water absorption coefficient, to quantitatively estimate $\ts(\freqsym)$ and $\sv(\freqsym)$. Methods to estimate these are not within the scope of this paper, but common practise is to use the conventional sphere backscatter calibration methodology \citep{demerCalibrationAcousticInstruments2015} slightly enhanced for broadband \citep{hobaekCharacterizationTargetSpheres2013,lavery2017}. We note that these methods do not provide an operational method to estimate $\teff$ or $\Phi(f)$, especially for ship-mounted transducers, and that empirical measurements of these parameters are necessary to fully calibrate both narrowband and broadband echosounders.

A set of equations and associated computer code for calculating calibrated, frequency-dependent, target strength and volume backscatter from broadband echosounder signals have been presented. This provides a resource for those interested in learning and further developing broadband processing techniques. The processing equations and methodology presented in this paper are similar to those implements in version 1.12.4 and earlier of the \ek{} software.
